Greenhouse gas emissions model - making scenarios

### 2021-11-13 Scenarios_report.Rmd
breed intervention - first complete draft

variables_to_update_posnorm<-c(paste0("milk_yield_",4:6),
                             "feed_prod_CO2",
                             "feed_trans_CO2",
                             paste0("live_weight_LW_",1:12),
                             paste0("weight_gain_WG_",1:12),
                             "DE_perc",
                             "perc_crude_protein_diet")

@@

### error estimates for variables
milk_yield_error<-0.25
feed_trans_error<-0.25
feed_prod_error<-0.25
live_weight_error<-0.1
weight_gain_error<-0.1
digestible_energy_error<-0.1
crude_protein_error<-0.1

@@

  # update feed production CO2 (with error)
   scenarios[which(scenarios$Variable == "feed_prod_CO2"),coln][1:2]<-
     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"feed_kgCO2"])*
                           c(1-feed_prod_error,1+feed_prod_error)
  # update feed transport CO2 (with error)
   scenarios[which(scenarios$Variable == "feed_trans_CO2"),coln][1:2]<-
     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"feed.transport_kgCO2"])*
                           c(1-feed_trans_error,1+feed_trans_error)
  # update milk yield for each cow type (animal type 4-6) (with error)
   for(typ in 4:6)
     if(sum(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ)>0)
       scenarios[which(scenarios$Variable==paste0("milk_yield_",typ)),coln][1:2]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "milk_yield"])*c(1-milk_yield_error,1+milk_yield_error)
  # update live weight and weight gain for each animal type (with error)
   for(typ in 1:12)
     if(sum(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ)>0)
       {scenarios[which(scenarios$Variable==paste0("live_weight_LW_",typ)),coln][1:2]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "live_weight_LW"])*c(1-live_weight_error,1+live_weight_error)
       scenarios[which(scenarios$Variable==paste0("weight_gain_WG_",typ)),coln][1:2]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "weight_gain_WG"])*c(1-weight_gain_error,1+weight_gain_error)
     }

@@

```{r uncertainty_package}
uncertainty::varscatter(in_var = in_var,
                        out_var = out_var,
                        xlab = "Total Milk Yield",
                        ylab = "Total emissions")
```
```{r varkernel}
uncertainty::varkernel(in_var = in_var, 
                       out_var = out_var,
                        xlab = "Total Milk Yield",
                        ylab = "Total emissions")
```

@@

## Breed

@@

```{r}
aggregate(Kenyaherds$milk_yield, by=list(Kenyaherds$breed,Kenyaherds$animal_type), FUN=mean)

cows<-Kenyaherds[which(Kenyaherds$animal_type %in% 4:6),]

@@

GHG_simulation_scenarios$x

@@

Kenyaherds[,"milk_change"]<-0
cows<-which(Kenyaherds$animal_type %in% 4:6)
Kenyaherds$milk_change[cows]<- sapply(cows,function(x) 1/milk_summary$Intervention_factor[milk_summary$Breed==Kenyaherds$breed[x]]  )

milk_change<-aggregate(Kenyaherds$milk_change[cows], by=list(Kenyaherds$QQNO[cows]), FUN=mean)

@@

for(i in 1:nrow(Kenyaherds))
  {Kenyaherds[i,"live_weight_change"]<-1/live_weights$Intervention_factor[which(live_weights$animal_type==Kenyaherds$animal_type[i]&
                                                                                live_weights$breed==Kenyaherds$breed[i])]
#   Kenyaherds[i,"mature_weight_change"]<-1/mature_weights$Intervention_factor[which(mature_weights$animal_type==Kenyaherds$animal_type[i]&
#                                                                                mature_weights$breed==Kenyaherds$breed[i])]
   Kenyaherds[i,"weight_gain_change"]<-1/weight_gains$Intervention_factor[which(weight_gains$animal_type==Kenyaherds$animal_type[i]&
                                                                                weight_gains$breed==Kenyaherds$breed[i])]
}

live_weight_agg<-aggregate(Kenyaherds$live_weight_change, by=list(Kenyaherds$QQNO,Kenyaherds$animal_type), FUN=mean)
#mature_weight_agg<-aggregate(Kenyaherds$mature_weight_change, by=list(Kenyaherds$QQNO,Kenyaherds$animal_type), FUN=mean)
weight_gain_agg<-aggregate(Kenyaherds$weight_gain_change, by=list(Kenyaherds$QQNO,Kenyaherds$animal_type), FUN=mean)

@@

for(i in 1:nrow(x_breed_intervention))
{if(rbinom(1,1,adoption_rate))
    x_breed_intervention[,colnames(breed_intervention)[2:ncol(breed_intervention)]]
  # implement changes in x file
}
  res<-run_model(x_breed_intervention,ghg_emissions)

@@

```{r}
uncertainty::varkernel(in_var = res$x$milk_yield_5, 
                       out_var = res$y$on_farm,
                        xlab = "Total Milk Yield",
                        ylab = "Total emissions")

@@

>>>>

variables_to_update_posnorm<-c(paste0("milk_yield_",4:6),
                             paste0("Cp_rate_",4:6),
                             "feed_prod_CO2",
                             "feed_trans_CO2",
                             paste0("live_weight_LW_",1:12),
                             paste0("weight_gain_WG_",1:12),
                             "DE_perc",
                             "perc_crude_protein_diet")

@@

### error estimates for variables
milk_yield_error<-0.275 # Migose et al. (2020, https://www.sciencedirect.com/science/article/abs/pii/S1871141319307371 )
feed_trans_error<-0.25 # Vellinga et al (2013, https://edepot.wur.nl/254098)
feed_prod_error<-0.25 # Vellinga et al (2013, https://edepot.wur.nl/254098)
live_weight_error<-0.145 # Goopy et al. (2018, https://www.publish.csiro.au/an/an16577 ) 
weight_gain_error<-0.145 # Goopy et al. (2018, https://www.publish.csiro.au/an/an16577 ) 
digestible_energy_error<-0.1
crude_protein_error<-0.1

@@

  # update feed production CO2 (with error)
   scenarios[which(scenarios$Variable == "feed_prod_CO2"),coln][1:2]<-
     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"feed_kgCO2"])*
                           c(1-feed_prod_error,1+feed_prod_error)
  # update feed transport CO2 (with error)
   scenarios[which(scenarios$Variable == "feed_trans_CO2"),coln][1:2]<-
     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"feed.transport_kgCO2"])*
                           c(1-feed_trans_error,1+feed_trans_error)
  # update milk yield for each cow type (animal type 4-6) (with error)
   for(typ in 4:6)
     if(sum(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ)>0)
       {scenarios[which(scenarios$Variable==paste0("milk_yield_",typ)),coln][1:2]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "milk_yield"])*c(1-milk_yield_error,1+milk_yield_error)
       scenarios[which(scenarios$Variable==paste0("CP_rate_",typ)),coln][1:2]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "Cp"]/0.1) *c(1,1)

     }
  # update live weight and weight gain for each animal type (with error)
   for(typ in 1:12)
     if(sum(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ)>0)
       {scenarios[which(scenarios$Variable==paste0("live_weight_LW_",typ)),coln][1:2]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "live_weight_LW"])*c(1-live_weight_error,1+live_weight_error)
       scenarios[which(scenarios$Variable==paste0("weight_gain_WG_",typ)),coln][1:2]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "weight_gain_WG"])*c(1-weight_gain_error,1+weight_gain_error)
     }

@@

```{r uncertainty_package}
uncertainty::varscatter(in_var = in_var,
                        out_var = out_var,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
```
```{r varkernel}
uncertainty::varkernel(in_var = in_var, 
                       out_var = out_var,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
```

@@

## Breed intervention

@@

```{r}
cows<-Kenyaherds[which(Kenyaherds$animal_type %in% 4:6),]

@@

@@

Kenyaherds_interventions<-Kenyaherds[,c("QQNO",
                                        "system",
                                        "animal_type",
                                        "breed",
                                        "live_weight_LW",
                                        "mature_weight_MW",
                                        "weight_gain_WG",
                                        "milk_yield",
                                        "dig_energy",
                                        "CP.")]

Kenyaherds_interventions[,"milk_change"]<-0
cows<-which(Kenyaherds_interventions$animal_type %in% 4:6)
Kenyaherds_interventions$milk_change[cows]<- sapply(cows,function(x) 1/milk_summary$Intervention_factor[milk_summary$Breed==Kenyaherds_interventions$breed[x]]  )

milk_change<-aggregate(Kenyaherds_interventions$milk_change[cows], by=list(Kenyaherds_interventions$QQNO[cows]), FUN=mean)

@@

for(i in 1:nrow(Kenyaherds_interventions))
  {Kenyaherds_interventions[i,"live_weight_change"]<-
    1/live_weights$Intervention_factor[which(live_weights$animal_type==Kenyaherds_interventions$animal_type[i]&
                                             live_weights$breed==Kenyaherds_interventions$breed[i])]
#   Kenyaherds_interventions[i,"mature_weight_change"]<-
#      1/mature_weights$Intervention_factor[which(mature_weights$animal_type==Kenyaherds_interventions$animal_type[i]&
#                                                 mature_weights$breed==Kenyaherds_interventions$breed[i])]
   Kenyaherds_interventions[i,"weight_gain_change"]<-
     1/weight_gains$Intervention_factor[which(weight_gains$animal_type==Kenyaherds_interventions$animal_type[i]&
                                              weight_gains$breed==Kenyaherds_interventions$breed[i])]
}

live_weight_agg<-aggregate(Kenyaherds_interventions$live_weight_change,
                           by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
#mature_weight_agg<-aggregate(Kenyaherds_interventions$mature_weight_change,
#by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
weight_gain_agg<-aggregate(Kenyaherds_interventions$weight_gain_change,
                           by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)

@@

for(i in 1:nrow(x_breed_intervention))
{if(rbinom(1,1,adoption_rate))
    x_breed_intervention[i,colnames(breed_intervention)[2:ncol(breed_intervention)]]<-
      x_breed_intervention[i,colnames(breed_intervention)[2:ncol(breed_intervention)]] * 
      breed_intervention[which(paste0("Farm_",breed_intervention$Farm)==x_breed_intervention$Scenario[i]),
                         2:ncol(breed_intervention)]
  # implement changes in x file
}
  breed_res<-run_model(x_breed_intervention,ghg_emissions)

  animals_per_run<-rowSums(breed_res$x[paste0("N_animal_type_",1:12)])

breed_milk_yield_per_head<-breed_res$y$milk_yield/animals_per_run
breed_total_emissions_per_head<-breed_res$y$on_farm/animals_per_run

@@

### Evaluate impact of breed intervention

Assumed adoption rate, based on all farms (even the ones that already have high-performance farms and therefore don't get an 'upgrade'): `r adoption_rate*100` %

```{r, fig.show="hold", out.width="50%", fig.cap="Breed intervention (left) vs. baseline (right)", message=FALSE}

uncertainty::varscatter(in_var = breed_milk_yield_per_head,
                        out_var = breed_total_emissions_per_head,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$milk_yield/animals_per_run,
                        out_var = GHG_simulation_scenarios$y$on_farm/animals_per_run,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernel(in_var = breed_milk_yield_per_head, 
                       out_var = breed_total_emissions_per_head,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$milk_yield/animals_per_run, 
                       out_var = GHG_simulation_scenarios$y$on_farm/animals_per_run,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernelslicerange(in_var = breed_res$y$milk_yield/animals_per_run, 
                                 out_var = breed_res$y$on_farm/animals_per_run, 
                                 min_in_var = round(min(breed_res$y$milk_yield/animals_per_run)),
                                 max_in_var = round(max(breed_res$y$milk_yield/animals_per_run)), 
                                 xlab_vars = "Total emissions given milk yield range")

uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$milk_yield/animals_per_run, 
                                 out_var = GHG_simulation_scenarios$y$on_farm/animals_per_run, 
                                 min_in_var = round(min(GHG_simulation_scenarios$y$milk_yield/animals_per_run)),
                                 max_in_var = round(max(GHG_simulation_scenarios$y$milk_yield/animals_per_run)), 
                                 xlab_vars = "Total emissions given milk yield range")

@@

```{r}
breed_change_in_CO2_eq_per_animal<-breed_res$y$on_farm/animals_per_run-
  GHG_simulation_scenarios$y$on_farm/animals_per_run

breed_change_in_milk_yield_per_animal<-breed_res$y$milk_yield/animals_per_run-
  GHG_simulation_scenarios$y$milk_yield/animals_per_run
```
